Factor Analysis on Citations, Using a Combined Latent and Logistic Regression Model

### 2019-10-01 Manuscript for submission/manuscript.tex
Update

%Astute readers might notice the slight difference of the first term in objective function between (\ref{eq:logistic10}) and (\ref{eq:Modified}).
For the convenience of theoretical investigation, we slightly modify the first term in the objective function summing over all $(i,j)$ pairs.
After scaling, due to symmetry of $X$,$L$, and $S$, the only difference between (\ref{eq:logistic10}) and (\ref{eq:Modified}) is in the inclusion of terms in diagonal pairs $(i,i),\forall i=1,\dots,n$.
%the first term in the objective function in (\ref{eq:Modified}) sums over all $(i,j)$ pairs.
This slight modification leads to neither theoretical consequence nor noticeable difference in practice.
Let ($\hat{\alpha},\widehat{L},\widehat{S}$) be the solution to (\ref{eq:Modified}), and ($\alpha^{*},L^{*},S^{*}$) be the ground truth, which governs the data generating process.
Let $\widehat{\Theta},\Theta^{*}$ be defined respectively as $\widehat{\Theta} = \hat{\alpha}\mathbbm{1}\mathbbm{1}^T+\widehat{L}+\widehat{S}$ and $\Theta^{*} = \alpha^{*}\mathbbm{1}\mathbbm{1}^T+L^{*}+S^{*}$.
And denote the error term for each parameter as $\widehat{\Delta}^{\Theta} = \widehat{\Theta}-\Theta^{*},
\hat{\Delta}^{\alpha} = \hat{\alpha}-\alpha^{*},
\widehat{\Delta}^L = \widehat{L}-L^{*},
\widehat{\Delta}^S = \widehat{S}-S^{*}.$
Throughout the discussion, let $P^{*}=\big\{\frac{\exp(\Theta_{ij}^{*})}{1+\exp(\Theta_{ij}^{*})}\big\}_{1 \leq i,j \leq n} \in \mathbb{R}^{n \times n}$.
We then impose several assumptions for theoretical guarantees of our estimator.

\begin{assumption}(\textbf{Strong convexity})  \label{Ass:1}
For any $\Theta \in \mathbb{R}^{n\times n}$, define the log-likelihood in (\ref{eq:Modified}), $h(\Theta) = -\frac{1}{n}\sum_{i,j} \big\{ X_{ij}\Theta_{ij} - \log(1+\exp(\Theta_{ij})) \big\}$.We assume that $h(\Theta)$ is $\tau$-strongly convex in a sense that lowest eigenvalue of Hessian matrix of the log-likelihood function is bounded away from zero ($\tau > 0$):
\[
\nabla^{2}h(\Theta) = \mbox{diag}\Big(\mbox{vec}\Big(\frac{\exp(\Theta)}{n(1+\exp(\Theta))^{2}}
\Big)\Big) \succcurlyeq \tau I_{n^{2} \times n^{2}}
\]
For any vector $a$, $\mbox{diag}(a)$ is the diagonal matrix with elements of $a$ on its diagonal. For any square matrix $A$ and $B$, we have $ A \succcurlyeq B $ if and only if matrix $A-B$ is positive semi-definite.
\end{assumption}

@@

With these assumptions, we present the behavior of non-asymptotic error bound of our estimator through the following theorem. In our result, we measure error using squared Frobenius norm summed across three matrices:
\[
    e^{2}(\hat{\alpha}\mathbbm{1}\mathbbm{1}^T,\widehat{L},\widehat{S}):=\|\hat{\Delta}^{\alpha}\mathbbm{1}\mathbbm{1}^T\|_{F}^{2} + \|\widehat{\Delta}^{L}\|_{F}^{2} + \|\widehat{\Delta}^{S}\|_{F}^{2}
\]

\begin{theorem} \label{Th:th1}
Under the assumptions \ref{Ass:1} and \ref{Ass:2},
if we solve the convex problem (\ref{eq:Modified}) with a pair of regularization parameter $(\delta,\gamma)$ satisfying
\begin{align} \label{eq:49}
\delta \geq 2\|\frac{1}{n}(X-P^{*})\|_{op} \quad and \quad \gamma \geq 2\|\frac{1}{n}(X-P^{*})\|_{\infty}+4\kappa\tau\bigg(\frac{Cn+1}{n} \bigg),
\end{align}
then there exist universal constants $c_{j}$, j = 1,2,3,  for all integers $k = 1,2,...,n$, and $s = 1,2,...,n^{2}$, we have the following upper bound on $e^{2}(\hat{\alpha}\mathbbm{1}\mathbbm{1}^T,\widehat{L},\widehat{S})$
\begin{align}
    e^{2}(\hat{\alpha}\mathbbm{1}\mathbbm{1}^T,\widehat{L},\widehat{S}) \leq
    \underbrace{c_{1}\frac{\delta^{2}}{\tau^{2}}}
    _{\mathcal{K}_{\alpha^*}} +
    \underbrace{c_{2}\frac{\delta^{2}}{\tau^{2}}
    \bigg\{k + \frac{\tau}{\delta}\sum_{j=k+1}^{n}\sigma_{j}(L^{*})\bigg\}}_{\mathcal{K}_{L^*}} +
    \underbrace{c_{3}\frac{\gamma^{2}}{\tau^{2}}
    \bigg\{s + \frac{\tau}{\gamma}\sum_{(i,j) \notin M}|S^*_{ij}|\bigg\}}_{\mathcal{K}_{S^*}},
\end{align}
where $M$ is an arbitrary subset of matrix indices of cardinality at most s.
\end{theorem}

@@

\section{ Numerical analysis and Applications}
\label{sec:Num-anal}
Section \ref{sec:Num-anal} is divided into two parts.
In subsection \ref{sec:Syn-dat}, we conduct an empirical study of measuring the performance of our proposed method with artificially specified graphical structures.
In subsection \ref{sec:citation_network}, we perform a real data analysis with citation network for statisticians.

@@

\subsection{Numerical experiments with synthetic data}
\label{sec:Syn-dat}
After an introduction on two synthetic scenarios that we want to explore, we describe three model selection criteria and four evaluation metrics for the selected model.
Subsequently, we elaborate experimental results from the synthetic networks and several interesting findings from those results.

@@

\begin{enumerate}
    \item We draw an intercept term $\alpha^*$ in the logistic regression model from the uniform distribution that is supported on [-11,-10].
        %In this way, we can make $\alpha^*$ have the least effects in creating edges in the network.

    \item Recall that the binary factor loading matrix $F^*$ encodes the relation between topics and papers
        (i.e., if $i$th paper studies about $k$th topic, we denote $F^*_{ki}=1$, otherwise $F^*_{ki}=0$).
        First, we simply set there exists $K$ topics in the network with $n$ papers, and each of them consists of $\frac{n}{K}$ papers.
        This can be expressed in $F^*$ as follows:
        \[
            F^* =
               \begin{bmatrix}
                 \smash[b]{\block{n/K}} \\
                 && \smash[b]{\block{n/K}} \\
                 &&&& \ddots \\
                 &&&&& \block{n/K}
                \end{bmatrix}
              \in R^{K \times n},
        \]
        where each row of $F^*$ has $\frac{n}{K}$ $1$s and each column has only one $1$.
        Then, we assume that $n_{l}$ papers share $l$ topics and $n_{m}$ papers share $m$ topics ($1 < l < m \leq K$).
        This can be represented in the $F^*$ in following steps:
        $1)$ Pick distinct $n_{l}$ indices randomly from $\{1,2,\dots,n\}$.
        We will denote the set of the indices as $\Omega_{n_l}$.
        $2)$ Choose $n_{m}$ indices from the set $\{1,2,\dots,n\}\setminus \Omega_{n_{l}}$ and denote the set of those indices as $\Omega_{n_{m}}$.
        $3)$ Make the columns of $F^*$ with corresponding indices in set $\Omega_{n_{l}} \cup \Omega_{n_{m}}$ zeros.
        $4)$ We use a notation $f^*_j$ to denote the $j$th column of the matrix $F^*$.
        Fill arbitrary $l$ entries of $f^*_j$ for $j \in \Omega_{n_l}$ with $1$s, and also fill arbitrary $m$ entries of $f^*_j$ for $j \in \Omega_{n_m}$ with $1$s.
        Lastly, we set $F^*=JF^*$ where $J=I_{n}-\frac{1}{n}\mathbbm{1}\mathbbm{1}^T$.

    \item Generate weight coefficients of the factors $D^*_{ii}$ from the uniform distribution that is supported on $[19,20]$ $ \forall 1 \leq i \leq K$.

    \item Remind that the positive entries of $S^*$ can characterize the links in the network which cannot be accounted by the common topics.
        In the spirit of this notion, we construct ad-hoc links in a way that they connect the clusters of papers with one topic.
        Specifically, we assume that there are $\frac{|S^*|}{_KC_{2}}$ edges between two clusters, where $|S^*|$ denotes the number of non-zero entries of the upper-triangular part of the matrix $S^*$.
        This can be implemented via following steps:
        1) We construct $K$ sets $C^*_1,\dots,C^*_K$ that are defined as follows:
        \begin{align*}
           &C^*_{1} \subseteq \big\{1, 2, \dots, \frac{n}{K} \big\} \setminus \big\{ \Omega_{n_l}
           \cup \Omega_{n_m} \big\} \\
           &C^*_{2} \subseteq \big\{\frac{n}{K}+1, \dots, \frac{2n}{K} \big\} \setminus \big\{ \Omega_{n_l} \cup \Omega_{n_m} \big\}  \\
           &\vdots \\
           &C^*_{K} \subseteq \big\{\frac{(K-1)n}{K}+1, \dots, n \big\} \setminus \big\{ \Omega_{n_l}
           \cup \Omega_{n_m} \big\}
        \end{align*}
        where each of them has arbitrary $\frac{|S^*|}{_KC_{2}}$ elements.
        2) Create a set $I_{S^*}$ whose elements are pairs of indices such that
        \[
            I_{S^*} =
            \bigg\{ (i_r,j_r) : i_r \in C^*_p, j_r \in C^*_q, 1 \leq p < q \leq K,   r = 1,2,\dots,\frac{|S^*|}{_KC_{2}} \bigg\},
        \]
        where we use $i_r$ to indicate $r$th element $i$ in the set.
        3) Draw $S^*_{ij} \sim \mbox{Unif}[19,20], \forall (i,j) \in I_{S^*}$.
        4) Lastly, make it symmetric by setting $S^*={S^*}^{T}$.
    \item Create an upper-triangular part of the adjacency matrix $X$ whose each entry $X_{ij}$ follows Bernoulli distribution.
        The distribution's parameter is parametrized by a probability, $P^{*}_{ij}=\frac{\exp (\alpha^{*}+f^{*T}_{i}D^{*}f^{*}_{j} +  S^{*}_{ij})}{1+\exp(\alpha^{*}+f^{*T}_{i}D^{*}f^{*}_{j} +  S^{*}_{ij})}$.
        After drawing all the entries of $X$ in the upper-triangular part, then make the matrix symmetric by setting $X=X^T$.
\end{enumerate}

@@

All six networks that are elaborated in scenarios $1$ and $2$ are visualized in Fig\ref{fig:figure1}.
Notice that the nodes that share the common topics are clustered,
and the cross clustered links are ad-hoc citations.
%ad-hoc links, which connect clusters of papers with $1$ topic, are colored in green.
%Readers can refer the R code provided in $\dots$, for more detailed descriptions on parametric settings for generating networks in Fig\ref{fig:figure1}.

@@

\subsubsection{Choosing the tuning parameters and evaluation criteria}
We observe that choosing a good pair of tuning parameters is an important yet challenging issue in our setting.
Here we present a heuristic procedure for choosing a good pair of tuning parameters $(\gamma,\delta)$.
Following the scree-plot approach in Ji and Jin \cite{ji2016coauthorship}, we plot the largest $15$ eigenvalues of the adjacency matrix $X$, and find an ``elbow'' point where the eigenvalues seem to level off.
An index of the point, which is to the left of this elbow point, is considered as the number of the topics embedded in the network.
(We will denote this number as $\widehat{K}^{\text{Scree}}$.)
We want to note the readers that the scree-plot analysis serves as a good guideline for determining the range of grids to search over.
With the estimate of the number of topics in the network in mind, we record the $\text{rank}(\widehat{L}^{\gamma,\delta})$ and $|\widehat{S}^{\gamma,\delta}|$ for each tuning parameter pair on a given grid.
We need to go through several iterations of this recording procedure to find a proper range of grid, in which we can get $\text{rank}(\widehat{L}^{\gamma,\delta})=\widehat{K}^{\text{Scree}}$
and $10^{-4}|X| \leq |\widehat{S}^{\gamma,\delta}| \leq 10^{-1}|X|$, via adjusting the range of grid for $\gamma$ and $\delta$ repeatedly.
Once we find a grid, which satisfies above constraints, we choose a pair of tuning parameters: %($\gamma^{\text{Heu}},\delta^{\text{Heu}}$):
\[
    (\gamma^{\text{Heu}},\delta^{\text{Heu}}):=
    \big\{(\gamma,\delta): \text{rank}(\widehat{L}^{\gamma,\delta})=\widehat{K}^{\text{Scree}},
    \quad \mbox{mode} |\widehat{S}^{\gamma,\delta}| \quad
    \text{s.t.} \quad 10^{-4}|X| \leq |\widehat{S}^{\gamma,\delta}| \leq 0.05|X| \big\}
\]
One might wonder how traditional model selection methods, such as the Bayes Information Criterion (BIC;\cite{schwarz1978estimating}) and the AIC, work where we recall that BIC and AIC are defined as follows:
\[
\mbox{BIC}(M) = -2  \mathbb{L}_n( \hat{\beta}(M)) + |M| \log \bigg(\frac{n(n-1)}{2}\bigg),
\]
and
\[
\mbox{AIC}(M) = -2  \mathbb{L}_n( \hat{\beta}(M)) + 2|M|.
\]
Here $M$ indicates the current model, which is implicitly understood that the model is obtained from certain tuning parameter pair $(\gamma,\delta)$.
We use $\mathbb{L}_n( \hat{\beta}(M))$ to denote the maximal log-likelihood for a given model $M$,
and $|M|$ is the number of free parameters in $M$, which is determined by the number of non-zeros in $\widehat{S}^{\gamma,\delta}$ and the low-rank matrix $\widehat{L}^{\gamma,\delta}$.
In detail, if we have $\text{rank}(\widehat{L}^{\gamma,\delta})=K$, we can establish the following
\[
|M| = \sum_{i < j}1_{\{S_{ij} \neq 0\}} + n K - \frac{K(K-1)}{2} + 1 ;
\]
since the number of free parameters in $\widehat{L}^{\gamma,\delta}$ is $K$ plus $nK - K(K+1)/2$, which is the number of free parameters in determining $K$ orth-normal vectors.
Additional $1$ in the last term is due to $\hat{\alpha}$.
We want to find a pair $(\gamma,\delta)$, which minimizes BIC($M$) and AIC($M$) as a function of $(\gamma,\delta)$ respectively, where we denote them as follows:
\[
    (\gamma^{BIC},\delta^{BIC}):= \mbox{arg min}_{\gamma, \delta}
    \mbox{BIC}(M),
    \quad
    (\gamma^{AIC},\delta^{AIC}):= \mbox{arg min}_{\gamma, \delta}
    \mbox{AIC}(M)
\]
We evaluate the models that are selected via our heuristic approach, BIC, and AIC by using following four evaluation metrics.

\begin{eqnarray*}
M_1 &=& \mathbbm{1}\big\{\mbox{rank}(\widehat{L}) = \mbox{rank}(L^{*})\big\}, \\
M_2 &=& \frac{\left|\big\{(i,j):i<j:S^{*}_{i,j} \neq0 \quad \& \quad \widehat{S}_{i,j} \neq 0 \big\}\right|}{\left|\big\{(i,j):i<j:S^{*}_{i,j}\neq0\big\}\right|}, \\
M_3 &=& \frac{\left|\big\{(i,j):i<j:S^{*}_{i,j} = 0 \quad \& \quad \widehat{S}_{i,j} \neq 0 \big\}\right|}{\left|\big\{(i,j):i<j:S^{*}_{i,j}=0\big\}\right|}, \\
M_4 &=& \frac{\left|\big\{\mbox{Misclassified Nodes}\big\}\right|}{n},
\end{eqnarray*}
where $M_1$ is a metric on whether the selected model recovers the true low rank structure of network,
$M_2$ evaluates the positive selection rate of the sparse ad-hoc structure in network,
and $M_3$ evaluates the false discovery rate of ad-hoc edges.
$M_4$ calculates the proportion of mis-classified nodes to the entire nodes in the network.
With properly selected tuning parameter, $M_1$ will be 1, $M_2$ will be close to 1, and $M_3$ and $M_4$ will get close to 0.
We present the evaluation results of the six cases via the four criteria, $M_1,M_2,M_3$, and $M_4$ in Table \ref{tab:table1}.

>>>>

%Astute readers might notice the slight difference of the first term in objective function between (\ref{eq:logistic10}) and (\ref{eq:Modified}).
For the convenience of theoretical investigation, we slightly modify the first term in the objective function summing over all $(i,j)$ pairs.
After scaling, due to symmetry of $X$,$L$, and $S$, the only difference between (\ref{eq:logistic10}) and (\ref{eq:Modified}) is in the inclusion of terms in diagonal pairs $(i,i),\forall i=1,\dots,n$.
%the first term in the objective function in (\ref{eq:Modified}) sums over all $(i,j)$ pairs.
This slight modification leads to neither theoretical consequence nor noticeable difference in practice.
Let ($\widehat{\alpha},\widehat{L},\widehat{S}$) be the solution to (\ref{eq:Modified}), and ($\alpha^{*},L^{*},S^{*}$) be the ground truth, which governs the data generating process.
Let $\widehat{\Theta}$ and $\Theta^{*}$ be defined respectively as $\widehat{\Theta} = \widehat{\alpha}\mathbbm{1}\mathbbm{1}^T+\widehat{L}+\widehat{S}$ and $\Theta^{*} = \alpha^{*}\mathbbm{1}\mathbbm{1}^T+L^{*}+S^{*}$.
And denote the error term for each parameter as $\widehat{\Delta}^{\Theta} = \widehat{\Theta}-\Theta^{*},
\widehat{\Delta}^{\alpha} = \widehat{\alpha}-\alpha^{*},
\widehat{\Delta}^L = \widehat{L}-L^{*},
\widehat{\Delta}^S = \widehat{S}-S^{*}.$
Throughout the discussion, let $P^{*}=\big\{\frac{\exp(\Theta_{ij}^{*})}{1+\exp(\Theta_{ij}^{*})}\big\}_{1 \leq i,j \leq n} \in \mathbb{R}^{n \times n}$.
We then impose several assumptions for theoretical guarantees of our estimator.

\begin{assumption}(\textbf{Strong convexity})  \label{Ass:1}
For any $\Theta \in \mathbb{R}^{n\times n}$, define the log-likelihood in (\ref{eq:Modified}), $h(\Theta) = -\frac{1}{n}\sum_{i,j} \big\{ X_{ij}\Theta_{ij} - \log(1+\exp(\Theta_{ij})) \big\}$.We assume that $h(\Theta)$ is $\tau$-strongly convex in a sense that lowest eigenvalue of Hessian matrix of the log-likelihood function is bounded away from zero ($\tau > 0$):
\[
\nabla^{2}h(\Theta) = \mbox{diag}\Big(\mbox{vec}\Big(\frac{\exp(\Theta)}{n(1+\exp(\Theta))^{2}}
\Big)\Big) \succcurlyeq \tau I_{n^{2} \times n^{2}}
\]
For any vector $a$, $\mbox{diag}(a)$ is the diagonal matrix with elements of $a$ on its diagonal.
For any matrix $B=[b_1,\dots,b_n]\in\mathbb{R}^{n \times n}$, $\mbox{vec}(B)\in\mathbb{R}^{n^2}$ is obtained by stacking $ b_1,\dots, b_{n}$ in order.
For any square matrix $A$ and $B$, we have $ A \succcurlyeq B $ if and only if matrix $A-B$ is positive semi-definite.
\end{assumption}

@@

With these assumptions, we present the behavior of non-asymptotic error bound of our estimator through the following theorem. In our result, we measure error using squared Frobenius norm summed across three matrices:
\[
    e^{2}(\widehat{\alpha}\mathbbm{1}\mathbbm{1}^T,\widehat{L},\widehat{S})
    :=\|\widehat{\Delta}^{\alpha}\mathbbm{1}\mathbbm{1}^T\|_{F}^{2} + \|\widehat{\Delta}^{L}\|_{F}^{2} + \|\widehat{\Delta}^{S}\|_{F}^{2}
\]

\begin{theorem} \label{Th:th1}
Under the assumptions \ref{Ass:1} and \ref{Ass:2},
if we solve the convex problem (\ref{eq:Modified}) with a pair of regularization parameter $(\delta,\gamma)$ satisfying
\begin{align} \label{eq:49}
\delta \geq 2\|\frac{1}{n}(X-P^{*})\|_{op} \quad and \quad \gamma \geq 2\|\frac{1}{n}(X-P^{*})\|_{\infty}+4\kappa\tau\bigg(\frac{Cn+1}{n} \bigg),
\end{align}
then there exist universal constants $c_{j}$, j = 1,2,3,  for all integers $k = 1,2,...,n$, and $s = 1,2,...,n^{2}$, and we have the following upper bound on $e^{2}(\widehat{\alpha}\mathbbm{1}\mathbbm{1}^T,\widehat{L},\widehat{S})$
\begin{align}
    e^{2}(\widehat{\alpha}\mathbbm{1}\mathbbm{1}^T,\widehat{L},\widehat{S}) \leq
    \underbrace{c_{1}\frac{\delta^{2}}{\tau^{2}}}
    _{\mathcal{K}_{\alpha^*}} +
    \underbrace{c_{2}\frac{\delta^{2}}{\tau^{2}}
    \bigg\{k + \frac{\tau}{\delta}\sum_{j=k+1}^{n}\sigma_{j}(L^{*})\bigg\}}_{\mathcal{K}_{L^*}} +
    \underbrace{c_{3}\frac{\gamma^{2}}{\tau^{2}}
    \bigg\{s + \frac{\tau}{\gamma}\sum_{(i,j) \notin M}|S^*_{ij}|\bigg\}}_{\mathcal{K}_{S^*}},
\end{align}
where $M$ is an arbitrary subset of matrix indices of cardinality at most s.
\end{theorem}

@@

\section{ Numerical analysis and Applications}
\label{sec:Num-anal}
Section \ref{sec:Num-anal} is divided into two parts.
In Section \ref{sec:Syn-dat}, we conduct an empirical study of our proposed method with artificial graphical structures.
In Section \ref{sec:citation_network}, we perform a real data analysis with citation network for statisticians.

@@

\subsection{Numerical experiments with synthetic data}
\label{sec:Syn-dat}
First, we introduce two synthetic scenarios that we want to explore.
Then, we describe three model selection criteria and four evaluation metrics for the selected model.
Subsequently, we elaborate experimental results from the synthetic networks and several interesting findings from those results.

@@

\begin{enumerate}
    \item We draw an intercept term $\alpha^*$ in the logistic regression model from the uniform distribution that is supported on [-11,-10].
        In this way, we can make $\alpha^*$ have the least effects in creating edges in the network.

    \item Recall that the binary factor loading matrix $F^*$ encodes the relation between topics and papers
        (i.e., if $i$th paper studies about $k$th topic, we denote $F^*_{ki}=1$, otherwise $F^*_{ki}=0$).
        First, we assume that there are $n$ papers in the network, and $K$ topics are embedded in the network. Each of them consists of $\frac{n}{K}$ papers.
        This can be expressed in $F^*$ as follows:
        \[
            F^* =
               \begin{bmatrix}
                 \smash[b]{\block{n/K}} \\
                 && \smash[b]{\block{n/K}} \\
                 &&&& \ddots \\
                 &&&&& \block{n/K}
                \end{bmatrix}
              \in R^{K \times n},
        \]
        where each row of $F^*$ has $\frac{n}{K}$ $1$'s and each column has only one $1$.
        Then, we assume that $n_{l}$ papers share $l$ topics and $n_{m}$ papers share $m$ topics ($1 < l < m \leq K$).
        This can be represented in the $F^*$ in following steps:
        $1)$ Pick distinct $n_{l}$ indices randomly from $\{1,2,\dots,n\}$.
        We will denote the set of the indices as $\Omega_{n_l}$.
        $2)$ Choose $n_{m}$ indices from the set $\{1,2,\dots,n\}\setminus \Omega_{n_{l}}$ and denote the set of those indices as $\Omega_{n_{m}}$.
        $3)$ Make the columns of $F^*$ with corresponding indices in set $\Omega_{n_{l}} \cup \Omega_{n_{m}}$ zeros.
        $4)$ We use a notation $f^*_j$ to denote the $j$th column of the matrix $F^*$.
        Fill arbitrary $l$ entries of $f^*_j$ for $j \in \Omega_{n_l}$ with $1$s, and also fill arbitrary $m$ entries of $f^*_j$ for $j \in \Omega_{n_m}$ with $1$s.
        Lastly, we set $F^*=JF^*$ where $J=I_{n}-\frac{1}{n}\mathbbm{1}\mathbbm{1}^T$.

    \item Generate the weight coefficients of the factors $D^*_{ii}$ from the uniform distribution that is supported on $[19,20]$ $ \forall 1 \leq i \leq K$.
        In this way, we can let the papers with same topic cluster together.

    \item Remind that the positive entries of $S^*$ can characterize the links in the network which cannot be accounted by the common topics.
        In the spirit of this notion, we construct ad-hoc links in a way that they connect the clusters of papers with one topic.
        Specifically, we assume that there are $\frac{|S^*|}{_KC_{2}}$ edges between two clusters, where $|S^*|$ denotes the number of non-zero entries of the upper-triangular part of the matrix $S^*$.
        This can be implemented via following steps:
        1) We construct $K$ sets $C^*_1,\dots,C^*_K$ that are defined as follows:
        \begin{align*}
           &C^*_{1} \subseteq \big\{1, 2, \dots, \frac{n}{K} \big\} \setminus \big\{ \Omega_{n_l}
           \cup \Omega_{n_m} \big\} \\
           &C^*_{2} \subseteq \big\{\frac{n}{K}+1, \dots, \frac{2n}{K} \big\} \setminus \big\{ \Omega_{n_l} \cup \Omega_{n_m} \big\}  \\
           &\vdots \\
           &C^*_{K} \subseteq \big\{\frac{(K-1)n}{K}+1, \dots, n \big\} \setminus \big\{ \Omega_{n_l}
           \cup \Omega_{n_m} \big\}
        \end{align*}
        where each of them has arbitrary $\frac{|S^*|}{_KC_{2}}$ elements.
        2) Create a set $I_{S^*}$ whose elements are pairs of indices such that
        \[
            I_{S^*} =
            \bigg\{ (i_r,j_r) : i_r \in C^*_p, j_r \in C^*_q, 1 \leq p < q \leq K,   r = 1,2,\dots,\frac{|S^*|}{_KC_{2}} \bigg\},
        \]
        where we use $i_r$ to indicate $r$th element $i$ in the set.
        3) Draw $S^*_{ij} \sim \mbox{Unif}[19,20], \forall (i,j) \in I_{S^*}$.
        4) Lastly, make it symmetric by setting $S^*_{jj}=S^*_{ij},\forall 1\leq i < j \leq n$.
    \item Create an upper-triangular part of the adjacency matrix $X$ whose each entry $X_{ij}$ follows Bernoulli distribution.
        The distribution's parameter is parametrized by a probability, $P^{*}_{ij}=\frac{\exp (\alpha^{*}+f^{*T}_{i}D^{*}f^{*}_{j} +  S^{*}_{ij})}{1+\exp(\alpha^{*}+f^{*T}_{i}D^{*}f^{*}_{j} +  S^{*}_{ij})}$.
        After drawing all the entries of $X$ in the upper-triangular part, then make the matrix symmetric by setting $X_{ji}=X_{ij},\forall 1 \leq i < j \leq n$.
\end{enumerate}

@@

All six networks that are elaborated in scenarios $1$ and $2$ are visualized in Fig.\ref{fig:figure1}.
Notice that the nodes that share the common topics are clustered,
and the cross clustered links are ad-hoc citations.
%ad-hoc links, which connect clusters of papers with $1$ topic, are colored in green.
%Readers can refer the R code provided in $\dots$, for more detailed descriptions on parametric settings for generating networks in Fig\ref{fig:figure1}.

@@

\subsubsection{Choosing the tuning parameters and evaluation criteria}
Choosing a good pair of tuning parameters is an important yet challenging issue in our setting.
Here we present a heuristic procedure for choosing a good pair of tuning parameters $(\gamma,\delta)$.
Following the scree-plot approach in Ji and Jin \cite{ji2016coauthorship}, we plot the largest $15$ eigenvalues of the adjacency matrix $X$, and find an ``elbow'' point where the eigenvalues seem to level off.
An index of the point, which is to the left of this elbow point, is considered as the number of the topics embedded in the network.
(We will denote this number as $\widehat{K}^{\text{Scree}}$.)
We want to note the readers that the scree-plot analysis serves as a good guideline for determining the range of grids to search over.
With the estimate of the number of topics in the network in mind, we record the $\text{rank}(\widehat{L}^{\gamma,\delta})$ and $|\widehat{S}^{\gamma,\delta}|$ for each tuning parameter pair on a given grid.
We need to go through several iterations of this recording procedure to find a proper range of grid, in which we can get $\text{rank}(\widehat{L}^{\gamma,\delta})=\widehat{K}^{\text{Scree}}$
and $10^{-4}|X| \leq |\widehat{S}^{\gamma,\delta}| \leq 10^{-1}|X|$, via adjusting the range of grid for $\gamma$ and $\delta$ repeatedly.
Once we find a grid, which satisfies above constraints, we choose a pair of tuning parameters: %($\gamma^{\text{Heu}},\delta^{\text{Heu}}$):
\[
    (\gamma^{\text{Heu}},\delta^{\text{Heu}}):=
    \big\{(\gamma,\delta): \text{rank}(\widehat{L}^{\gamma,\delta})=\widehat{K}^{\text{Scree}},
    \mbox{mode} |\widehat{S}^{\gamma,\delta}|: \mbox{subject to} 10^{-4}|X| \leq |\widehat{S}^{\gamma,\delta}| \leq 10^{-1}|X| \big\}
\]
One might wonder how traditional model selection methods, such as the Bayes Information Criterion (BIC;\cite{schwarz1978estimating}) and the AIC, work where we recall that BIC and AIC are defined as follows:
\[
\mbox{BIC}(M) = -2  \mathbb{L}_n( \hat{\beta}(M)) + |M| \log \bigg(\frac{n(n-1)}{2}\bigg),
\]
and
\[
\mbox{AIC}(M) = -2  \mathbb{L}_n( \hat{\beta}(M)) + 2|M|.
\]
Here $M$ indicates the current model, which is implicitly understood that the model is obtained from certain tuning parameter pair $(\gamma,\delta)$.
We use $\mathbb{L}_n( \hat{\beta}(M))$ to denote the maximal log-likelihood for a given model $M$,
and $|M|$ is the number of free parameters in $M$, which is determined by the number of non-zeros in $\widehat{S}^{\gamma,\delta}$ and the low-rank matrix $\widehat{L}^{\gamma,\delta}$.
In detail, if we have $\text{rank}(\widehat{L}^{\gamma,\delta})=K$, we can establish the following
\[
|M| = \sum_{i < j}1_{\{S_{ij} \neq 0\}} + n K - \frac{K(K-1)}{2} + 1 ;
\]
since the number of free parameters in $\widehat{L}^{\gamma,\delta}$ is $K$ plus $nK - K(K+1)/2$, which is the number of free parameters in determining $K$ orth-normal vectors.
Additional $1$ in the last term is due to $\hat{\alpha}$.
We want to find a pair $(\gamma,\delta)$, which minimizes BIC($M$) and AIC($M$) as a function of $(\gamma,\delta)$ respectively, where we denote them as follows:
\[
    (\gamma^{BIC},\delta^{BIC}):= \mbox{arg min}_{\gamma, \delta}
    \mbox{BIC}(M),
    \quad
    (\gamma^{AIC},\delta^{AIC}):= \mbox{arg min}_{\gamma, \delta}
    \mbox{AIC}(M)
\]
We evaluate the models that are selected via our heuristic approach, BIC, and AIC by using following four evaluation metrics.

\begin{eqnarray*}
M_1 &=& \mathbbm{1}\big\{\mbox{rank}(\widehat{L}) = \mbox{rank}(L^{*})\big\}, \\
M_2 &=& \frac{\left|\big\{(i,j):i<j:S^{*}_{i,j} \neq0 \quad \& \quad \widehat{S}_{i,j} \neq 0 \big\}\right|}{\left|\big\{(i,j):i<j:S^{*}_{i,j}\neq0\big\}\right|}, \\
M_3 &=& \frac{\left|\big\{(i,j):i<j:S^{*}_{i,j} = 0 \quad \& \quad \widehat{S}_{i,j} \neq 0 \big\}\right|}{\left|\big\{(i,j):i<j:S^{*}_{i,j}=0\big\}\right|}, \\
M_4 &=& \frac{\left|\big\{\mbox{Misclassified Nodes}\big\}\right|}{n},
\end{eqnarray*}
where $M_1$ is a metric on whether the selected model recovers the true low rank structure of network,
$M_2$ evaluates the positive selection rate of the sparse ad-hoc structure in network,
and $M_3$ evaluates the false discovery rate of ad-hoc edges.
$M_4$ calculates the proportion of mis-classified nodes to the entire nodes in the network.
With properly selected tuning parameter, $M_1$ will be 1, $M_2$ will be close to 1, and $M_3$ and $M_4$ will get close to 0.
We present the evaluation results of the six cases via the four criteria, $M_1,M_2,M_3$, and $M_4$ in Table.\ref{tab:table1}.

\subsubsection{Several Observations}
%Given a network data $X$, after fitting the model with a proper pair of tuning parameters, $(\gamma,\delta)$, we need to determine whether node $i$ belongs to $k$th topic or not.
%We apply a simple $k$-means clustering on the fitted $L$ matrix's $K$ eigenvectors.
%For the illustration of performance of our fitted model after applying $k$-means algorithm on three synthetic scenarios, in Fig\ref{fig:figure1}, we plot the coordinates of each rows of first two eigen-vectors on $2$D plane.
%Furthermore, we find one interesting phenomenon.
%Coordinates of each rows from first two eigenvectors of estimated $\widehat{L}$ matrix in our model characterize the clustering behavior of embedded topics reasonably well.
%Our experience shows that with only first two eigenvectors of $\widehat{L}$, we can also cluster the nodes comparably well.
\begin{enumerate}
    \item \textbf{Model Selection.}
    First and foremost, choosing a good pair of tuning parameters is critical when it comes to making a good statistical inference on data.
    As presented in Table.\ref{tab:table1}, both BIC and AIC, which are well known for their model selection consistency in asymptotic setting, appear to under-estimate either the number of topics or the number of ad-hoc links in the networks in our synthetic settings. This may be caused by the fact that these traditional methods take the sample size into account, and therefore penalizes the model complexity too harshly.
    In the heuristic approach, scree-plot plays an important role when it comes to recovering the number of topics, and this strategy leads us to good model selection results for all the six cases considered in two scenarios. (See Fig.\ref{fig:figure2})
